### 2020-02-07 ncov_lancet_ID.tex
stuff

\usepackage{natbib}
\usepackage[superscript,biblabel]{cite}
\setcitestyle{numbers} 
\setcitestyle{square}
\bibliographystyle{unsrt}
\date{\today}

@@

\begin{flushleft}{
	\Large
	\textbf\newline{
		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
	}
}
\newline
\\
Sang Woo Park BSc\textsuperscript{1,*}
Benjamin M.\ Bolker PhD\textsuperscript{2,3,4}
David Champredon PhD\textsuperscript{5}
David J.\,D.\ Earn PhD\textsuperscript{3,4}
Michael Li PhD\textsuperscript{2}
Joshua S.\ Weitz PhD\textsuperscript{6, 7}
Bryan T.\ Grenfell PhD\textsuperscript{1,8,9}
Jonathan Dushoff PhD\textsuperscript{2,3,4,*}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
\\
\textbf{6} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{7} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{8} Division of International Epidemiology and Population Studies, Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\textbf{9} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\bigskip

@@

A novel coronavirus (2019-nCoV) has recently emerged as a global threat.
As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
We present a framework for comparing and combining different estimates of \Ro across a wide range of models.

@@

We reviewed 7 studies that were published online between January 23--26, 2020.
We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
We assess early estimates of \Ro for the 2019-nCoV outbreak using the pooled estimates.

\subsection*{Findings}

We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
In particular, estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.

\subsection*{Interpretation}

@@

\subsection*{Funding}

BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).

\section*{Research in context}

\subsection*{Evidence before this study}

% 7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
% Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.

This section should include a description of all the evidence that the authors considered before undertaking this study. Authors should briefly state: the sources (databases, journal or book reference lists, etc) searched; the criteria used to include or exclude studies (including the exact start and end dates of the search), which should not be limited to English language publications; the search terms used; the quality (risk of bias) of that evidence; and the pooled estimate derived from meta-analysis of the evidence, if appropriate

\subsection*{Added value of this study}

Authors should describe here how their findings add value to the existing evidence

\subsection*{Implications of all the available evidence}

Authors should state the implications for practice or policy and future research of their study combined with existing evidence.Research in context panels should not contain references; key studies mentioned here should be referenced in the main text.

@@

As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups when the estimation methods and their underlying assumptions vary widely.

@@

\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & Sources \\
\hline
Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
\hline
Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
\hline
Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
\hline
Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
\hline
Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
\hline
Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
\hline
Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
}
\end{table}

@@

\section*{Contributors}

@@

\bibliography{ncov}

>>>>

\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{unsrtnat}
\date{\today}

@@

\begin{flushleft}{
	\Large
	\textbf\newline{
		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1,*}
Benjamin M.\ Bolker\textsuperscript{2,3,4}
David Champredon\textsuperscript{5}
David J.\,D.\ Earn\textsuperscript{3,4}
Michael Li\textsuperscript{2}
Joshua S.\ Weitz\textsuperscript{6, 7}
Bryan T.\ Grenfell\textsuperscript{1,8,9}
Jonathan Dushoff\textsuperscript{2,3,4,*}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
\\
\textbf{6} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{7} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{8} Division of International Epidemiology and Population Studies, Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\textbf{9} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\bigskip

@@

A novel coronavirus (2019-nCoV) has recently emerged as a global threat.
As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.

\subsection*{Aim}

We aimed to develop a framework for comparing and combining different estimates of \Ro across a wide range of models.

@@

We reviewed 7 model-based analyses of the 2019-nCoV outbreak that were published online between January 23--26, 2020.
We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.

\subsection*{Results}

We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
Estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.

\subsection*{Conclusion}

@@

\section*{Keywords}

Basic reproductive number, 2019-nCoV

\section*{Funding}

BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR). The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\section*{Contribution}

@@

As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints\cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups when the estimation methods and their underlying assumptions vary widely.

@@

\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
\hline
Study 1 & 1.5--3.5 & 10 & 1 & Bedford \textit{et al.} \cite{bedfordncov} \\
\hline
Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & Imai \textit{et al.} \cite{imaincov} \\
\hline
Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & Liu \textit{et al.} \cite{liuncov} \\
\hline
Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & Read \textit{et al.} \cite{readncov} \\
\hline
Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & Riou and Althaus \cite{riouncov} \\
\hline
Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & Zhao \textit{et al.} \cite{zhaoncov} \\
\hline
Study 7 & 2.0--3.1 & 6--10 & 0 & Majumder and Mandl \cite{majumderncov} \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
}
\end{table}

@@

@@

\bibliography{ncov_abbr}
